% This file was converted to LaTeX by Writer2LaTeX ver. 0.4
% see http://www.hj-gym.dk/~hj/writer2latex for more info
\documentclass[12pt,twoside]{article}
\usepackage[ascii]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsfonts,textcomp}
\usepackage{color}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue}
% List styles
\newcommand\liststyleWWviiiNumvi{%
\renewcommand\labelitemi{[F0A7?]}
\renewcommand\labelitemii{o}
\renewcommand\labelitemiii{[F0A7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
\newcommand\liststyleLi{%
\renewcommand\labelitemi{[E00A?]}
\renewcommand\labelitemii{[25CB?]}
\renewcommand\labelitemiii{[25A0?]}
\renewcommand\labelitemiv{[25CF?]}
}
\newcommand\liststyleWWviiiNumiii{%
\renewcommand\labelitemi{[F0A7?]}
\renewcommand\labelitemii{o}
\renewcommand\labelitemiii{[F0A7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
\newcommand\liststyleWWviiiNumv{%
\renewcommand\labelitemi{[F0A7?]}
\renewcommand\labelitemii{o}
\renewcommand\labelitemiii{[F0A7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
\newcommand\liststyleWWviiiNumix{%
\renewcommand\labelitemi{[F0A7?]}
\renewcommand\labelitemii{o}
\renewcommand\labelitemiii{[F0A7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
\newcommand\liststyleWWviiiNumiv{%
\renewcommand\labelitemi{[F0A7?]}
\renewcommand\labelitemii{o}
\renewcommand\labelitemiii{[F0A7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
% Pages styles (master pages)
\makeatletter
\newcommand\ps@Standard{%
\renewcommand\@oddhead{}%
\renewcommand\@evenhead{}%
\renewcommand\@oddfoot{}%
\renewcommand\@evenfoot{}%
\setlength\paperwidth{8.5in}\setlength\paperheight{11in}\setlength\voffset{-1in}\setlength\hoffset{-1in}\setlength\topmargin{0.7874in}\setlength\headheight{12pt}\setlength\headsep{0cm}\setlength\footskip{12pt+0cm}\setlength\textheight{11in-0.7874in-0.9846in-0cm-12pt-0cm-12pt}\setlength\oddsidemargin{0.9846in}\setlength\textwidth{8.5in-0.9846in-0.9846in}
\renewcommand\thepage{\arabic{page}}
\setlength{\skip\footins}{0.0398in}\renewcommand\footnoterule{\vspace*{-0.0071in}\noindent\textcolor{black}{\rule{0.25\columnwidth}{0.0071in}}\vspace*{0.0398in}}
}
\makeatother
\pagestyle{Standard}
\title{Niels Joubert}
\begin{document}
\clearpage\pagestyle{Standard}
{\raggedleft\selectlanguage{english}\sffamily
Niels Joubert
\par}

{\raggedleft\selectlanguage{english}
\textsf{Berkley, CA 94704 \ \ {\textbullet} \ \ (650) 823{}-1662
\ \ {\textbullet} \ \ njoubert.blogspot.com \ \ {\textbullet}
\ }\href{mailto:niels@berkeley.edu}{\textsf{\textcolor{blue}{niels@berkeley.edu}}}
\par}

{\centering\selectlanguage{english}
[2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?][2017?]
\par}


\bigskip

{\selectlanguage{english}\sffamily\bfseries
Education}


\bigskip

\liststyleWWviiiNumvi
\begin{enumerate}
\item {\selectlanguage{english}\sffamily\bfseries
University of California at Berkeley }
\end{enumerate}
{\selectlanguage{english}\sffamily\mdseries
 B.S. in Electrical Engineering and Computer Sciences, expected
graduation: 2009.}

{\selectlanguage{english}\sffamily\mdseries
 Cumulative GPA: 4.0}

\liststyleWWviiiNumvi
\begin{enumerate}
\item {\selectlanguage{english}\sffamily
\textbf{Los Altos High School, Grade 12: }\textmd{(Los Altos, CA),
Graduated Valedictorian, GPA 4.0}}
\item {\selectlanguage{english}
\textsf{\textbf{Mountain View Adult Education Center: (Mountain View,
CA) }}\textsf{\textmd{concurrent enrollment while attending Los Altos
High School.}}}
\item {\selectlanguage{english}
\textsf{\textbf{Paul Roos Gymnasium, Grade 8 {--} 11:}}\textsf{
(Stellenbosch, South Africa)}}
\end{enumerate}

\bigskip

{\selectlanguage{english}\sffamily\bfseries
Skills and Experience}


\bigskip

\liststyleWWviiiNumvi
\begin{enumerate}
\item {\selectlanguage{english}\sffamily\bfseries
Web Development \textmd{using RubyOnRails, MySQL, HTML, CSS, JavaScript,
AJAX, XML{}-RPC, REST and screen{}-scraping.}}
\item {\selectlanguage{english}\sffamily\bfseries
Software Development~\textmd{using Java5, Junit testing and UNIX build
environment}}
\item {\selectlanguage{english}\sffamily\bfseries
Operating System/Tools\textmd{ comfortable in Windows, Linux and Solaris
development environments using Emacs and Eclipse.}}
\end{enumerate}

\bigskip

{\selectlanguage{english}\sffamily\bfseries
Projects}

\liststyleLi
\begin{enumerate}
\item {\selectlanguage{english}\sffamily\mdseries
Logical Constraint problem solver to analyze word problems.}
\item {\selectlanguage{english}\sffamily\mdseries
Particle Simulator using Quadtree data structure to enable O(log4(n))
lookup of objects, handling upwards of 10 000 datapoints in real{}-time
graphical simulation.}
\item {\selectlanguage{english}\sffamily\mdseries
Network{}-enabled board game using Game Tree algorithms to implement
quasi{}-intelligent computer player and remote playing capabilities
over TCP networks.}
\item {\selectlanguage{english}\sffamily\mdseries
Web{}-based portal and organizational site for Berkeley students,
involving interfaces to Berkeley{\textquotesingle}s academic resources
and web{}-based }
\end{enumerate}

\bigskip

{\selectlanguage{english}\sffamily\bfseries
Employment history}


\bigskip

\liststyleWWviiiNumvi
\begin{enumerate}
\item {\selectlanguage{english}
\textsf{\textbf{The Caf\'e Table.com, Spring 2006 {--} present:
}}\textsf{\textmd{Co{}-founder of web{}-based startup focusing on
integration of online technologies and portal services to present a
unified, customizable interface for Berkeley Student{\textquotesingle}s
academic and social lives. (http://thecafetable.com/)}}}
\item {\selectlanguage{english}
\textsf{\textbf{UC Berkeley New Student Services, Jan 2006 {--} August
2006}}\textsf{: Cal Student Orientation counselor, responsible for
facilitating and leading groups of new freshman and transfer students
through several two{}-day programs to acquaint them with the academic,
professional and social atmosphere of the University of California at
Berkeley.}}
\item {\selectlanguage{english}
\textsf{\textbf{TIDE Productions, Jan 2003 {--} Jan 2006}}\textsf{:
Involved in 5 successful DVD productions, repeatedly contracted by the
}\textsf{\textit{Stellenbosch Youth Orchestra,}}\textsf{ and produced a
DVD that was recently used for the successful application to the
position of Conductor at the Vancouver Conservatoire.}}
\end{enumerate}

\bigskip


\bigskip


\bigskip

{\selectlanguage{english}\sffamily\bfseries
Activities on and off campus}

\liststyleWWviiiNumiii
\begin{enumerate}
\item {\selectlanguage{english}\sffamily\bfseries
Web Developer}
\item {\selectlanguage{english}\sffamily\bfseries
Cal Linux User Group Member}
\item {\selectlanguage{english}
\textsf{\textbf{Musician:}}\textsf{ an accomplished violinist and bass
guitarist.}}
\item {\selectlanguage{english}
\textsf{\textbf{Radio Amateur: }}\textsf{\textmd{Licensed South African
radio amateur, }}\textsf{member of the American Radio Relay League.}}
\item {\selectlanguage{english}\sffamily\bfseries
Photographer and Cinematographer}


\bigskip

{\selectlanguage{english}\sffamily\bfseries
Service \& Leadership Experience}
\end{enumerate}
\liststyleWWviiiNumv
\begin{enumerate}
\item {\selectlanguage{english}
\textsf{\textbf{Volunteer webmaster \& team leader, Paul Roos Gymnasium:
}}\textsf{volunteered as webmaster at Paul Roos Gymnasium in 2003,
became team leader January 2004. Received Pro Merito award for
exceptional service, increased traffic 10 fold.}}
\item {\selectlanguage{english}
\textsf{\textbf{SunSTEP engineering for youth, 2000 {--} 2004:}}\textsf{
Alumnus of SunSTEP program, designed to create opportunities for
formerly disadvantaged youth to participate in electrical engineering
projects. Lab Assistantship and general member of SunSTEP during all 4
years.}}
\item {\selectlanguage{english}
\textsf{\textbf{BARK Amateur Radio Club, Presentation host \& technology
implementer, 2001 {--} 2003:}}\textsf{ Introduced technology advances
through lectures and public demonstrations. Successful implementation
and demonstration of said technologies bought about repeat lectures at
various clubs and employment of techniques in emergency communication
systems.}}
\end{enumerate}
\liststyleWWviiiNumix
\begin{enumerate}
\item {\selectlanguage{english}
\textsf{\textbf{Cal Student Orientation counselor:}}\textsf{ Group
leader and facilitator for incoming students, networking with 45 other
counselors and 200+ incoming students.}}
\end{enumerate}

\bigskip

{\selectlanguage{english}\sffamily\bfseries
Honors, Awards, Recognition}

\liststyleWWviiiNumiv
\begin{enumerate}
\item {\selectlanguage{english}\sffamily
Yahoo! University Hack Day winner at UC Berkeley.}
\item {\selectlanguage{english}\sffamily
Golden Key Club nominee as a student in the top 5\% of Berkeley.}
\item {\selectlanguage{english}\sffamily
Chancellor's Honors for outstanding academic achievement in the first
semester at the University of California at Berkeley.}
\item {\selectlanguage{english}\sffamily
Twice{}-received Richard G. Bell Memorial Scholarship for academic
achievement, 2005}
\item {\selectlanguage{english}\sffamily
National Dean's List member.}
\item {\selectlanguage{english}\sffamily
California Mathematics Council prize for Excellence in Mathematics,
2005}
\item {\selectlanguage{english}\sffamily
Bill Alheid award for number one scholar, Paul Roos Gymnasium, 2004}
\item {\selectlanguage{english}\sffamily
Gold Medal in Expo for Young Scientists, Electronical Engineering and
General Engineering prizes, 2002}
\end{enumerate}
\end{document}
